---
title: "COVID vaccination and myocarditis and pericarditis"
author: '[Fatemeh Torabi](mailto:fatemeh.torabi@swansea.ac.uk)'
date: "Last edited `r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '2'
    code_folding: hide
  word_document:
    toc: yes
    toc_depth: '2'
toc-title: Contents
always_allow_html: yes
header-includes: null
---

```{r global_options, include=FALSE}
library(knitr)
setwd("P:/torabif/workspace/CCU0002-03")
knitr::opts_chunk$set(fig.path = 'figures/', fig.align = 'center', echo = FALSE, warning = FALSE, message = FALSE)

pkgs <- c(
    "assertr",
    "beepr",
    "broom",
    "dplyr",
    "data.table",
    "dtplyr",
    "forcats",
    "ggplot2",
    "ggthemes",
    "knitr",
    "kableExtra",
    "mice",
    "janitor",
    "lubridate",
    "qs",
    "rmarkdown",
    "sailr",
    "scales",
    "stringr",
    "readr",
    "survival",
    "tableone",
    "tidyr",
    "RODBC",
    "Cairo",
    "lattice",
    "getopt"
)

for (pkg in pkgs) {
    suppressWarnings(
        suppressPackageStartupMessages(
            library(pkg, character.only = TRUE)
        )
    )
    message("\t", pkg, sep = "")
}
```

```{r "loginSctn", eval=TRUE, echo = FALSE }
source("S:/0911 - Utilising routine data and machine learning techniques to discover new multi/FT/R_login/login_box.r");
login = getLogin('torabif');
sql = odbcConnect('PR_SAIL',login[1],login[2]);
login = 0
```
<!-------------------------------------------------------------------------------------------->
\newpage
\tableofcontents
\newpage

## Introduction

There has been reports on possible association between the second dose of
COVID-19 vaccination (Pfizer-BioNTech and Moderna) with with myocarditis or 
myopericarditis. 
The patients affected were mostly males, younger than 50 years old and
had an illness of chest pain associated with fever and with a rise in cardiac 
specific troponins that later fully recovered. In some patients, cardiac 
magnetic resonance imaging (MRI) demonstrated typical signs of myocarditis.
Whilst this illness appears self-limiting, it is important to understand its 
frequency because these vaccines are currently the backbone of the COVID-19 
control program in younger people.

In addition, it is important to clarify whether the risk of events seen after
the second vaccination dose is similar or different to the risk after the 
first vaccination dose.

Myocarditis and pericarditis are rare conditions. We will conduct this study 
through pooled analyses of linked healthcare data from the populations of 
England and Wales (total populations of around 57 million and 3 million respectively). 
The very large population size is needed to detect sufficient numbers of these 
rare health conditions for analyses to be statistically robust. 

In addition, the linked health data come from mutiple sources (e.g. primary care,
secondary care, death registries, COVID testing and vaccination data), providing
comprehensive information for defining the exposure of interest (vaccination), the outcome
of interest (myocarditis or pericarditis), key characteristics of the population
(e.g., age, sex, ethnicity) and other important comorbidities (e.g. prior COVID infection) 

## Research Hypotheses

There is a higher risk of myocarditis or pericarditis after COVID-19 vaccination than before or
without vaccination, that varies with vaccine type and by first or second vaccination dose


## Timeline

* observation period: 2020-12-08 to 2021-06-30

## Phenotype list

The following list of phenotypes was provided to the project

```{r phenorg, eval=TRUE, echo=FALSE}

org.phen <- sqlQuery(sql,"
      SELECT * FROM (
    	SELECT DISTINCT name, code, DESC FROM SAILWWMCCV.PHEN_ICD10_MYOCARDITIS 
    	UNION 
    	SELECT DISTINCT name, code, DESC FROM SAILWWMCCV.PHEN_ICD10_PERICARDITIS
    	)	ORDER BY code;
        ")

#events <- format(events,big.mark=',')
kable(org.phen)%>%
  kable_paper("hover", full_width=F)

```

Codes were matched to the following diagnosis codes in Patient Episode Dataset for Wales (PEDW)

```{r phenpedw, eval=TRUE, echo=FALSE}

pedw.phen <- sqlQuery(sql,"
      SELECT DISTINCT ALT_CODE , DESCRIPTION FROM SAILUKHDV.ICD10_CODES_AND_TITLES_AND_METADATA icatam 
      WHERE ALT_CODE IN (SELECT DISTINCT DIAG_CD_1234 FROM sailwWMCCV.ccu0002_3_pedw_outcome)
      ORDER BY ALT_CODE;
                      ")
kable(pedw.phen)%>%
  kable_paper("hover", full_width=F)

```


## Cohort eligibility

**Please note:** That the <5 counts that can be seen in the 'Events Diff' column are referring to records have been removed due to either incomplete or erroneous information. This means that these are not a privacy risk.

```{r summary, echo = FALSE, message = FALSE}
d_analysis <- qread("data/cohort_raw.qs")

read_csv("results/t_sample_selection_summary.csv", col_types = cols()) %>%
kable(
    digits = 1,
    format.args = list(big.mark = ","),
    col.names = c(
        "Step",
        "Sample criteria",
        "n",
        "Diff",
        "(%)",
        "n",
        "Diff",
        "(%)"
    )
) %>%
kable_styling(
    bootstrap_options = "striped",
    full_width = FALSE
) %>%
add_header_above(c(
    " "                    = 2,
    "Individuals"          = 3,
    "Events"               = 3
))
```


Total number of cases in pre and post vaccinated periods

#### **Table 1:** Baseline characteristics for the cases and controls of the study cohort.

```{r baseline, echo = FALSE, message = FALSE}
library(qs)
d_analysis <- qread("data/analysis_cohort.qs")
#d_analysis$has_outcome[is.na(d_analysis$has_outcome)] <- 0

d_analysis[d_analysis$cov_dose1_age < 40,"age_cat"]="< 40";
d_analysis[d_analysis$cov_dose1_age >= 40 & d_analysis$cov_dose1_age < 69,"age_cat"]="4069";
d_analysis[d_analysis$cov_dose1_age >= 70,"age_cat"]="70";

xvar <- c(
    "cov_dose1_age",
    "age_cat",
    "cov_dose1_sex",
    "cov_dose1_ethnicity",
    "cov_dose1_deprivation",
    "vaccination_dose1_product",
    "vaccination_dose2_product"
    )

d_analysis$age_cat <- as.factor(d_analysis$age_cat )
d_analysis$cov_dose1_sex <- as.factor(d_analysis$cov_dose1_sex)
d_analysis$cov_dose1_ethnicity <- as.factor(d_analysis$cov_dose1_ethnicity)
d_analysis$cov_dose1_deprivation <- as.factor(d_analysis$cov_dose1_deprivation)
d_analysis$vaccination_dose1_product <- as.factor(d_analysis$vaccination_dose1_product)
d_analysis$vaccination_dose2_product <- as.factor(d_analysis$vaccination_dose2_product)

table_one <-
    d_analysis %>%
    CreateTableOne(
        vars = xvar,
        data = .,
##        strata = "has_outcome",
        test = FALSE
    )
library(tableone)

tab1<- as.data.frame(print(table_one, printToggle = FALSE, test=TRUE));


#A font package that allows bringing in Windows fonts. This was done so that we can use unicode characters.
#install.packages("extrafont",repo="http://cran.rstudio.com");
#library(extrafont)

library(pander)

name=as.character(c("n", 
                    "Age mean (sd)",
                    "Age category",
                    "< 40",
                    "40-69",
                    "70+",
                    "Sex",
                    "Ethnicity",
                    "Asian",
                    "Black",
                    "Mixed",
                    "Other",
                    "White",
                    "Deprivation WIMD quintile",
                    "  1 (most deprived)",
                    "  2","  3","  4",
                    "  5 (least deprived)", 
                    "First Vaccine",
                    "AstraZeneca",
                    "Moderna",
                    "Pfizer-BioNTech",
                    "Second Vaccine",
                    "AstraZeneca",
                    "Moderna",
                    "Pfizer-BioNTech"));

table1 = cbind(name, tab1);
colnames(table1) <- c("characteristics", "Overall")
#print using pander

panderOptions('table.alignment.default', function(table1) ifelse(apply(table1, 2, is.numeric), 'right', 'left'));
pander(table1, row.names=FALSE)

```


#### **Table 2:** The following table shows the number of events observed in pre and week1_2 and week3_23 post vaccination hazard ratio of myocarditis or pericarditis post vaccination. Please note that we don't provide at any of the tables total count of events broken down by vaccine type in overall post vacccination period and therefore the masked ('<5') counts for week1_2 and week3_23 can take the full range up to 5. 


```{r, echo = FALSE, message = FALSE}

cnt_vac1_az <-      read_csv("results/tbl_event_count_VAC1_any_myo_or_pericarditis_all_vac_az.csv",  col_types = cols())
cnt_vac1_pf<-       read_csv("results/tbl_event_count_VAC1_any_myo_or_pericarditis_all_vac_pf.csv", col_types = cols())
cnt_vac2_az<-       read_csv("results/tbl_event_count_VAC2_any_myo_or_pericarditis_all_vac_az.csv", col_types = cols())
cnt_vac2_pf <-      read_csv("results/tbl_event_count_VAC2_any_myo_or_pericarditis_all_vac_pf.csv", col_types = cols())

cnt_vac1_az <- cnt_vac1_az%>%
                mutate(vacdose="Dose1 - ChAdOx1")
cnt_vac1_pf <- cnt_vac1_az%>%
                mutate(vacdose="Dose1 - BNT162b2")
cnt_vac2_az <- cnt_vac2_az%>%
                mutate(vacdose="Dose2 - ChAdOx1")
cnt_vac2_pf <- cnt_vac2_az%>%
                mutate(vacdose="Dose2 - BNT162b2")
            

#model coef:

coef_vac1_az<-    read_csv("results/tbl_hr_+ weeks + agegroup + sex_VAC1_any_myo_or_pericarditis_all_vac_az.csv",               col_types = cols())
coef_vac1_pf <-   read_csv("results/tbl_hr_+ weeks + agegroup + sex_VAC1_any_myo_or_pericarditis_all_vac_pf.csv", col_types = cols())
coef_vac2_az <-   read_csv("results/tbl_hr_+ weeks + agegroup + sex_VAC2_any_myo_or_pericarditis_all_vac_az.csv", col_types = cols())
coef_vac2_pf<-    read_csv("results/tbl_hr_+ weeks + agegroup + sex_VAC2_any_myo_or_pericarditis_all_vac_pf.csv",      col_types = cols())

coef_vac1_az<- coef_vac1_az %>% 
                mutate(vacdose="Dose1 - ChAdOx1")
coef_vac1_pf<- coef_vac1_pf %>% 
                mutate(vacdose="Dose1 - BNT162b2")
coef_vac2_az<- coef_vac2_az %>% 
                mutate(vacdose="Dose2 - ChAdOx1")
coef_vac2_pf<- coef_vac2_pf %>% 
                mutate(vacdose="Dose2 - BNT162b2")


desc <-
    bind_rows(
        cnt_vac1_az,
        cnt_vac2_az,
        cnt_vac1_pf,
        cnt_vac2_pf
      )%>%
      dplyr::select(vacdose,expo_week,events_total)
desc$events_total[desc$events_total <5] <- "<5"
# pretty table -----------------------------------------------------------------

coef <-
    bind_rows(
    coef_vac1_az,
    coef_vac1_pf,
    coef_vac2_az,
    coef_vac2_pf
    ) %>%
    rename(expo_week=term )%>%
    dplyr::select(vacdose,expo_week,estimate,conf.low,conf.high)

coef$estimate <- round(coef$estimate,digits=2)
coef$conf.low <- round(coef$conf.low,digits=2)
coef$conf.high <- round(coef$conf.high,digits=2)
coef<- coef %>% 
  mutate(
        HR    =format(round(estimate, 2), nsmall=2),
        ci.95 = str_c("(",conf.low,"to",conf.high,")")
         )

desc_coef <-
    desc %>%
    full_join(coef, by = c("vacdose", "expo_week")) 

coef_table <- desc_coef %>% 
                  filter(expo_week %in% c("pre expo","week1_2","week3_23")) %>% 
                  dplyr::select(vacdose,expo_week,events_total,HR,ci.95)

panderOptions('table.alignment.default', function(coef_table) ifelse(apply(coef_table, 2, is.numeric), 'right', 'left'));
pander(coef_table, row.names=FALSE)
```

### Error bars for hazard ratio of events **post first vaccination**

```{r, echo = FALSE, message = FALSE}
library(ggthemes)
d1<- desc_coef %>% 
  filter(vacdose=="Dose1 - ChAdOx1" | vacdose=="Dose1 - BNT162b2") %>% 
  filter(expo_week %in% c("week1_2","week3_23"))

ggplot(d1, aes(x=expo_week, y=estimate, group=vacdose, color=vacdose))+
  geom_line(alpha= 0.4, size=1.5)+
  geom_point()+
  geom_errorbar(aes(ymin=conf.low, ymax=conf.high), width=0.2)+
    geom_hline(yintercept = 1, linetype='dotted',col="blue")+
  theme_classic()
#  scale_colour_manual(values=c("#B33951","#304C89"))

```


### Error bars for hazard ratio of events **post second vaccination**

```{r, echo = FALSE, message = FALSE}

d2<- desc_coef %>% 
  filter(vacdose=="Dose2 - ChAdOx1" | vacdose=="Dose2 - BNT162b2") %>% 
  filter(expo_week %in% c("week1_2","week3_23"))

ggplot(d2, aes(x=expo_week, y=estimate, group=vacdose, color=vacdose))+
  geom_line(alpha= 0.4, size=1.5)+
  geom_point()+
  geom_errorbar(aes(ymin=conf.low, ymax=conf.high), width=0.2)+
  geom_hline(yintercept = 1, linetype='dotted',col="blue")+
  theme_classic()

```
